---
title: "Fertility"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

To analyze outlier cases, we are going to use a dataset with fertility rates by country, and other fields such as per capita income.

```{r library, echo=FALSE}
library(tidyverse)
```

```{r fertility}
fertility <- read.csv("tasaFertilidad2019vsGPD.csv")
fertility %>% glimpse()
```

## Visual inspection

The first method that we will program is visual inspection, for which we will use the ggplot library.

```{r, echo=FALSE}
library(ggplot2)

ggplot(fertility, aes(x = TasaFertilidad, y = IngresoPerCapita)) +
  geom_point() + # esto genera un grafico de dispersion de TasaFertilidad vs IngresoPerCapita
  ggtitle("Fertility Rate vs Per Capita Income (2019)") + 
  xlab("Average number of children per woman") + 
  ylab("Average income per capita, in thousands of US$") + 
  theme_bw() +
  theme(axis.title = element_text(size = 15),
        axis.text = element_text(size = 10, color="black"),
        plot.title = element_text(size = 20, face="bold", hjust=0.5),
        text = element_text(size=25))
```

Anomalies can be seen in the graph, but it is a subjective method and dependent on each analyst.

## Grubb's univariate test

To use an analytical method, we will apply the Grubb test, which is implemented in the outliers library.

```{r, echo=FALSE}
library(outliers)

# We calculate the grub test on the PerCapita Income variable
dataOutlier <- grubbs.test(fertility$IngresoPerCapita, two.sided=F)
dataOutlier

# We calculate the grub test on the variable Fertility Rate
dataOutlier <- grubbs.test(fertility$TasaFertilidad, two.sided=F)
dataOutlier
```

This method has the limitation that the number of atypical data to be reviewed must be predefined. In this case, it validates that the maximum data of both variables have atypical behavior.

## Distance methods

A multivariate alternative to the detection of atypical cases is the distance method. We will calculate the Mahalanobis distance to consider the covariances present in the data.

```{r, echo=FALSE}
# calculate the distances using the mahalanobis function, available in base R
distM <- mahalanobis(fertility[,c(4,5)],colMeans(fertility[,c(4,5)]),cov(fertility[,c(4,5)]))
 
# visualize the distribution of distances
ggplot(distM %>% as_tibble(), aes(x=distM)) +
  geom_density() +
  theme_bw() + 
  theme(text=element_text(size=25))

# visualize the scatter plot.
ggplot(fertility, aes(x = TasaFertilidad, y = IngresoPerCapita, color = distM)) +
  scale_color_gradient(low="blue", high="red") +
  geom_point(size=1, alpha=0.8) +
  ggtitle("Fertility Rate vs Per Capita Income (2019)") + 
  xlab("Average number of children per woman") + 
  ylab("Average income per capita, in thousands of US$") + 
  theme_bw() +
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 10, color="black"),
        text=element_text(size=25),
        plot.title = element_text(size = 20, face="bold", hjust=0.5))
```

It can be seen that those points furthest from the central cloud have the greatest distance values, which can be seen in red.

## Nearest neighbors method

This method is based on identifying the neighbors with the shortest distance to each point, so the same mahalanobis distance can be used, but the exclusion criterion is more elaborate.

```{r}
library(distances)
library(dbscan)

# Calculate mahalanobis distances, using another method available in the distances library
tempDist <- distances(fertility[,c(4,5)],normalize = "mahalanobize") %>% as.dist()

# Preserve the 4 minor distances
temp <- kNNdist(tempDist, 4, all=T)
temp <- temp[,4] # preserve the maximum of the 4 close distances

# Visualize the density of these maximum distances
ggplot(temp %>% as_tibble(), aes(x=temp)) +
  geom_density() +
  theme_bw() +
  theme(text=element_text(size=25))

# visualize the scatter plot, colored according to maximum distance of knn
ggplot(fertility, aes(x=TasaFertilidad,y=IngresoPerCapita,color=temp)) +
  geom_point(size=1,alpha=0.8)+theme_bw()+
  scale_color_gradient(low="blue",high="red")+
  ggtitle("Fertility Rate vs Per Capita Income (2019)") +  
  theme_bw() +
  theme(plot.title = element_text(size = 20, face="bold", hjust=0.5),
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 10, color="black")) +
  xlab("Average number of children per woman") + 
  ylab("Average income per capita, in thousands of US$")
```

This analysis allows us to identify atypical data that are isolated but within the ranges of values of variables.

## Density method

To execute the density method, we will calculate the local outlier factor, or LOF, implemented in the dbscan library.

```{r, warning = FALSE, message=FALSE}
library(dbscan)

## Calculate the lof from mahalanobis distances calculated in the previous step (tempDist)
outlierScores = dbscan::lof(tempDist, k=5)

# Visualize the density of the abnormal scores
ggplot(outlierScores %>% as_tibble(), aes(x= outlierScores)) + 
  geom_density() + 
  theme_bw() + 
  theme(text=element_text(size=18))

# generate subset of data with high outlier scores
subData=fertility[outlierScores>3,]

# visualize dispersion, colored according to LOF
ggplot(fertility, aes(x = TasaFertilidad, y = IngresoPerCapita, color = outlierScores)) +  
  geom_point(size=1,alpha=0.8) +
  geom_text(data = subData, label = subData$Abrev, color = "blue") +  # visualize the texts (Abrev) of atypical values
  scale_color_gradient(low="blue", high="red")+
  ggtitle("Fertility Rate vs Per Capita Income (2019)") +  
  xlab("Average number of children per woman") + 
  ylab("Average income per capita, in thousands of US$")
```

According to this analysis, the GQ, IQ and NE points correspond to atypical values.
