---
title: "Churn_Tree"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For the Decision Trees chapter we are going to continue using Churn data, or customer churn. Before loading the data, we invoke the Tidyverse library. After loading the data, we cleaned it up a bit and took a look at it.

```{r, message= FALSE, warning = FALSE}
#tidymodels
library(tidyverse)

data <- read_csv("Churn_Modelling.csv") %>% 
  mutate(is_female = ifelse(Gender == "Female",1,0),
         Exited = as.factor(Exited)) %>% 
        select(-RowNumber, -Surname, -Geography, -Gender, -CustomerId) %>% 
  relocate(Exited)
 
data %>% glimpse()
```

We generate the same dataset from the previous class but with a single line of code.

Now, to implement the model we are going to use a library called tidymodels, which allows us to unify different Machine Learning libraries present in R. We also load the discrim library, which has complementary classification models to tidymodels.

The first step we do is to separate the data into training and test sets, where tidymodels has the initial_split function.

```{r, message = FALSE, warning=FALSE}
library(tidymodels)
library(discrim) 

data_split <- initial_split(data, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)

nrow(test_data)
train_data %>% nrow()
```

Next, we create the "recipe" of the model, which consists of the "black box" relationship between the input variables and the output variables. In this case, the recipe will be to model Exited based on all the variables present in the data set.

```{r}
receta <- 
  recipe(Exited ~ ., data = train_data) 

receta
```

Now if we create the model, where we will use a decision tree with 5 decision layers, and a minimum number of entities per leaf (pruning) of 10. The library used to calculate this model will be rpart, which comes preloaded in the packages we are using. With this step we only define the model, we still calculate it.

```{r}
modelo <-
  decision_tree(tree_depth = 5, min_n = 10) %>% 
  set_engine("rpart") %>% 
  set_mode("classification")

modelo
```

Now we fit the model, calculate its predictions and calculate the AUC value

```{r}
fitea <- function(mod){
  
  modelo_fit <- 
  workflow() %>% 
  add_model(mod) %>% 
  add_recipe(receta) %>% 
  fit(data = train_data)

model_pred <- 
  predict(modelo_fit, test_data, type = "prob") %>% 
  bind_cols(test_data) 

return(model_pred %>% 
  roc_auc(truth = Exited, .pred_0))
}

fitea(modelo)
```

Now we will see the magic of tidymodels, we will make a comparison with other models, such as the logistic regression model, naïve Bayes or Knn. For this, the only thing we need to change is the model, since the recipe is the same, and the validation flow is also the same. Therefore we can use the function we created above to evaluate the different models and compare them.

```{r, warning=FALSE}
modelo_rl <- 
  logistic_reg() %>% 
  set_engine("glm")

fitea(modelo_rl)

modelo_nb <-
  naive_Bayes(smoothness = .8) %>%
  set_engine("naivebayes")

fitea(modelo_nb)

modelo_knn <-
  nearest_neighbor(neighbors = 5) %>% 
  set_engine("kknn") %>% 
  set_mode("classification")

fitea(modelo_knn)
```

We can see that in this case the Naive Bayes model obtains the best results when classifying with an AUC of .84.