---
title: "datasauRus"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Analysis of the data set to summarize its main characteristics, using statistical and visual methods.

Goals:

- Discover the underlying structure of the data.

- Identify relevant variables.

- Detect outliers and anomalies.

- Validate assumptions.

- Generate hypotheses from the data.

## Exploratory Data Analysis (EDA)

We are going to explore a data set called datasauRus, which consists of the combination of 13 data sets with x and y coordinates.


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(datasauRus)
library(proxy)
```

```{r}
summary(datasaurus_dozen)
```

### Data statistics

We compute a variety of statistics and similarity metrics for each unique category in the "datasaurus_dozen" dataset and present a summary of these results using the glimpse() and summary() functions. Analysis is performed to explore and understand the characteristics of the data based on the "dataset" variable.

```{r}
stats <- datasaurus_dozen %>% 
            group_by(dataset) %>% 
            summarize(
              mean_x    = mean(x),
              mean_y    = mean(y),
              std_dev_x = sd(x),
              std_dev_y = sd(y),
              corr_pears  = cor(x, y, method = "pearson"),
              corr_spear  = cor(x, y, method = "spearman"),
              corr_kendall  = cor(x, y, method = "kendall"),
              simil_cos = simil(list(x,y), method = "cosine") %>% as.numeric(),
              simil_jac = simil(list(x,y), method = "Jaccard") %>% as.numeric(),
              simil_sm = simil(list(x,y), method = "simple matching") %>% as.numeric(),
              simil_kul = simil(list(x,y), method = "Kulczynski1") %>% as.numeric(),
              dist_euc = dist(list(x,y), method = "Euclidean") %>% as.numeric(),   
              dist_manh = dist(list(x,y), method = "Manhattan") %>% as.numeric(),  
              dist_sup = dist(list(x,y), method = "supremum") %>% as.numeric(),   
              median_x    = median(x),
              median_y    = median(y),
              CV_x = sd(x) / mean(x),
              CV_y = sd(y) / mean(y),
              max_x = max(x),
              max_y = max(y)
            )

stats %>% glimpse()

summary(pr_DB)
```

### Data Visualization

#### Dot plot

```{r}
ggplot(datasaurus_dozen, aes(x=x, y=y, colour=dataset)) +
    geom_point() +
    theme_void() +
    theme(legend.position = "none") +
    facet_wrap(~dataset, ncol=3)
```

#### Histogram

```{r}
ggplot(datasaurus_dozen, aes(x=x,colour=dataset)) +
    geom_histogram(binwidth = 2) +
    theme_void() +
    theme(legend.position = "none") +
    facet_wrap(~dataset, ncol=3)
```

#### Density

```{r}
ggplot(datasaurus_dozen, aes(x=x,colour=dataset))+
    geom_density()+
    theme_void()+
    theme(legend.position = "none")+
    facet_wrap(~dataset, ncol=3)
```

#### Boxplot

```{r}
ggplot(datasaurus_dozen, aes(x=x, colour=dataset))+
    geom_boxplot()+
    theme_void()+
    theme(legend.position = "none")+
    facet_wrap(~dataset, ncol=3)
```

#### Violin

```{r}
ggplot(datasaurus_dozen, aes(x=x, y=y, colour=dataset))+
    geom_violin()+
    theme_void()+
    theme(legend.position = "none")+
    facet_wrap(~dataset, ncol=3)
```

#### Boxplot_2

```{r}
ggplot(datasaurus_dozen, aes(x=x, y=dataset, colour=dataset)) +
    geom_point() +
    geom_boxplot() +
    theme_void() +
    theme(legend.position = "none")
```